---
title: "Introduction to geneference"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{geneference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

geneference aims to ...

In this vignette, we show an example of how one might go about conducting a simulation study relying solely on the functionalities implemented in geneference.

geneference contains functions for .. blabla .. the idea being to do as much in R as possible with the user worrying as little as possible about PLINK ..  
the general workflow is illustrated on the below diagram:



# Simulation
geneference .. several simulation functions.. all of these assume model see vignette.

We can ... see the framework directly in our R session:

```{r}
#this is sim_test
```
blabla simple and small.. blabla for larger, we need to work outside R(mention plink, write to disk)

We use `sim_random_family` to blabla... 
```{r, eval=FALSE}
#this is call to large dataset with random family size
```
.. we now have .. files in our directory .. (show this somehow?) .. to save space we chose .bed, this means that if we try to open the file it is gibberish...


# Analysis
we want to assign gwax, ltfh pheno
let's say we want to investigate the benefit of having data on more siblings. instead of spending time and space simulating a completely new dataset, we can simply limit the assigners to only use certain number of siblings

show that we don't need to know max number of siblings

call calculate cov - it actually tells us how many families with at least x siblings

call calculate error types, use pipe to get more rows
comment on general performance
## Visualisation
Examples of plots.
