---
title: "Documentation for our Gibbs sampling process"
description: >
  This vignette provides some details on the Gibbs sampler we use as part of
  computing the LT-FH phenotype.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Gibbs sampling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "../man/figures/gibbs-",
  out.width = "100%"
)
library(ggplot2)
```

This document is at the moment just a test. Once the Gibbs functions have been implemented to the package, we can begin doing stuff here.
I will probably speak with Rasmus tomorrow to try some stuff out.
For now, I just want to check that I can run a code chunk generating a plot without including the code, and that the figure is saved as expected.

```{r plottest, echo=FALSE}
x <- c(1, 2, 3, 4)
y <- c(5, 10, 15, 20)
mydata <- data.frame(x, y)
ggplot(mydata, aes(x, y))+
  geom_point()+
  theme_light()+
  NULL
```
What a beautiful plot that is.